---
title: "SLE712 assessment part 1 and part 2"
author: "Kinhgsley Isaac Ogamanya"
date: "18/05/2020"
output: html_document
theme: cosmo
---

##Question1: Read in the file, making the gene accession numbers the row names. Show a table of values for the first six genes.
```{r,download}
download.file("https://raw.githubusercontent.com/markziemann/SLE712_files/master/bioinfo_asst3_part1_files/gene_expression.tsv",
              destfile = "gene_expression.tsv")
```
I use the download.file command to import files from GitHub 
And then i read in R making sure the first columns is set as the row names and headers are recongnized 
so as to show a table of values for the first six genes, and then proceed into using the head command to check if the gene has been imported properly, by showing a table of values for the first six genes and also checking the data structure.
```{r,import_read}
gene <- read.table("gene_expression.tsv",header = TRUE , stringsAsFactors = FALSE , row.names = 1)
head(gene)
str(gene)
```

##Question2: make a new column which is the mean of the other columns. Show a table of values for the first six genes
```{r,chunk1}
gene$mean <- rowMeans(gene)
head(gene)
str(gene)
```
In this question2, i use the rowMeans command to create the mean of the new columns 
and then using the head command to check if the gene has been imported properly 
by showing a table of values for the first six genes and also checking the data structure


##Question3: List the 10 genes with the highest mean expression
```{r,chunk2}
head(gene[order(-gene$mean), ], 10)
```
In question 3, i use the order command and square brackets to reorder the gene files 
and providing their mean and then using the head command to check 
and list genes with the highest mean expression

##Question4: Determine the number of genes with a mean <10
```{r,chunk3}
filtered <- subset(gene, mean < 10)
nrow(filtered)
```
in question 4, i use the subset command to determine the genes with a mean <10
and then filtered to get the total number of genes with a mean <10

##Question5: Make a histogram plot of the mean values in png format and paste it into your report
```{r,plot1}
hist(gene$mean)
```
I created a histogram showing the frequency against the gene$mean

##Question6: Import this csv file into an R object. What are the column names?
```{r,chunk4}
download.file("https://raw.githubusercontent.com/markziemann/SLE712_files/master/bioinfo_asst3_part1_files/growth_data.csv",
              destfile = "growth_data.csv")
```
Using the download.file command to import files from GitHub 
And then i read in R using the `read.csv()`to obtain the dataset and also making sure the header are recongnized and seting the `stringsAsFcators = FALSE`and then proceed into using the head command to check if the gene has been imported properly, by showing a table of values for the first six genes and also checking the data structure, then finally use the `colnames` command to list the column names
```{r,chunk5}
tree <- read.csv("growth_data.csv", header = TRUE, stringsAsFactors = FALSE)
head(tree)
str(tree)
colnames(tree)
```

##Question7: Calculate the mean and standard deviation of tree circumference at the start and end of the study at both sites.
```{r,chunk6,echo=TRUE}
ne <- subset(tree,Site=="northeast")
head(ne)
tail(ne)
str(ne)
mean(ne$Circumf_2004_cm)
mean(ne$Circumf_2019_cm)
sd(ne$Circumf_2004_cm)
sd(ne$Circumf_2019_cm)

#repeat same procedure for the southwest 
sw <- subset(tree,Site=="southwest")
head(sw)
tail(sw)
str(sw)
mean(sw$Circumf_2004_cm)
mean(sw$Circumf_2019_cm)
sd(sw$Circumf_2004_cm)
sd(sw$Circumf_2019_cm)
```
In question 7, I first created two objects which are data frame for northeast and southwest, and then using the subset command to subset the tree growth obejcts so as to get the data that corresponds to both the northeast and southwest and then using the head and tail command to properly import and obtain the first and last 6 of the tree growths for both sites, making sure the structures are checked, and then proceed to calculate and obtain the mean and standard deviation of tree circumference at the start and end of the study for both sites using the mean and sd command

##Question8: make a boxplot of tree circumference at the start and end of the study at both sites
```{r,plot2}
boxplot(ne$Circumf_2004_cm,ne$Circumf_2019_cm,sw$Circumf_2004_cm,sw$Circumf_2019_cm,
        names =c("ne2004","ne2019","sw2004","sw2019"),ylab="Circumference (cm)" ,
        main="Tree growth at both plantation sites")
```
In question 8, i use the boxplot command to plot a chart showing the tree circumference at the start and end of the study for both sites

##Question9: Calculate the mean growth over the past 10 years at each site
```{r,chunk7,echo=TRUE}
ne$growth <-ne$Circumf_2019_cm-ne$Circumf_2009_cm
head(ne)
mean(ne$growth)

#repeat same procedure for the southwest
sw$growth <-sw$Circumf_2019_cm-sw$Circumf_2009_cm
head(sw)
mean(sw$growth)
```
Using the downloaded subset, i substract the past year which is 2009 from the recent year which is 2019 and using the head command to check if imported properly, and then using the mean command to calculate the mean growth over the past 10 years, repeated the steps for both sites.

##Question 10: use the t.test and wilcox.test functions to estimate the p-value and show that the 10 year growth is different at the two sites.
```{r,chunk8}
t.test(sw$growth,ne$growth)
wilcox.test(sw$growth,ne$growth)
```
i use the `t.test` and `wilcox.test` command to determine the difference between the two sites

