---
title: "SLE712 assessment part 1"
author: "Kinhgsley Isaac Ogamanya"
date: "18/05/2020"
output: html_document
theme: cosmo
---

##Question1: 
Read in the file, making the gene accession numbers the row names. 
Show a table of values for the first six genes.
```{r,download}
download.file("https://raw.githubusercontent.com/markziemann/SLE712_files/master/bioinfo_asst3_part1_files/gene_expression.tsv",
              destfile = "gene_expression.tsv")
```
I use the download.file command to import files from GitHub 
And then i read in R making sure the first columns is set as the row names and headers are recongnized 
so as to show a table of values for the first six genes, and then proceed into using the head command to check if the gene has been imported properly, by showing a table of values for the first six genes and also checking the data structure.
```{r,import_read}
gene <- read.table("gene_expression.tsv",header = TRUE , stringsAsFactors = FALSE , row.names = 1)
head(gene)
str(gene)
```

##Question2: 
make a new column which is the mean of the other columns. Show a table of values for the first six genes
```{r,chunk1}
gene$mean <- rowMeans(gene)
head(gene)
str(gene)
```
In this question2, i use the rowMeans command to create the mean of the new columns 
and then using the head command to check if the gene has been imported properly 
by showing a table of values for the first six genes and also checking the data structure


##Question3: 
List the 10 genes with the highest mean expression
```{r,chunk2}
head(gene[order(-gene$mean), ], 10)
```
In question 3, i use the order command and square brackets to reorder the gene files 
and providing their mean and then using the head command to check 
and list genes with the highest mean expression

##Question4: 
Determine the number of genes with a mean <10
```{r,chunk3}
filtered <- subset(gene, mean < 10)
nrow(filtered)
```
in question 4, i use the subset command to determine the genes with a mean <10
and then filtered to get the total number of genes with a mean <10

##Question5: 
Make a histogram plot of the mean values in png format and paste it into your report
```{r,plot1}
hist(gene$mean)
```
I created a histogram showing the frequency against the gene$mean